---
title: "Analysis for manuscript submitted to JASA-EL"
author: "Mats E. Nilsson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette describes analyses of the ALEXIS 108 data as presented in a manuscript first submitted to JASA-Express Letters 2018-02-23, and then resubmitted 2018-04-28. The analysis below refer to the revised manuscript. 

## R Libraries
Two R-libraries are used:

+ **alexis108** contains the data. It can be installed from github using the R library **devtools**, or you could just download the datafile `oskar_ild_mntraining.txt` from github: github.com/stamnosslin/alexis108/tree/master/data  

+ **rstan** is used for fitting psychometric functions to data using Hamiltonian Monte Carlo estimation.

+ **rethinking** is used for summarizing the posterior, using chainmode() for teh mode and HPDI() for 95 % credibility interval.

```{r message = FALSE}
# devtools::install_github("stamnosslin/alexis108")  # Installs alexis108 
library(alexis108)
library(rstan)
library(rethinking)
```
<br>
<br>

## Get data

```{r}
# Lateralization experiment, stored in data fram d
d <- data108
d$pc <- 1 * (d$correct == d$answer)  # pc = 1 if answer was correct, 0 if not
```

<br>
<br>

## Period during which training was conducted
Training was conducted in 60 separate days distributed over a period of this many days:
```{r}
first <- as.Date(d$date[1],'%Y-%m-%d')
last <- as.Date(d$date[nrow(d)],'%Y-%m-%d')
difftime(last, first, units = "days")
```
<br>
<br>

## Number of trials per 10 day period

```{r}
ndays <- function(ici) {
   dfun <- function(startday) {
     tau <- ici
     days <- seq(startday, startday + 9)
     length(d$pc[d$tau == tau & d$day %in% days])
   }
   
   sapply(1:51, dfun)
}

taus <- sort(unique(d$tau))  # Stimulus conditions (tau = ici)
ntrials <- sapply(taus, ndays)
summary(apply(ntrials, 2, mean))

```

<br>
<br>

## Function to convert LLR to ILD for the ICI = 0 condition
```{r}
llr_to_ild <- function(llr) {
  lmax <- 20 * log10 (10^(0/20) + 10^(llr/20))
  lmin <- 20 * log10 (10^(0/20) + 10^((llr-10)/20))
  lmax-lmin
}

# Example given in the Method section
llrs <- c(10, 0, -10, -20)  
round(sapply(llrs, llr_to_ild),1)

# LLR at threshold for ICI = 0 ms was -15 dB for days 1-10, and -17 dB LLR for days 51-60. This calculates the corresponding ILDs
round(sapply(c(-15.057090, -16.840977), llr_to_ild), 1) 
```

<br>
<br>

## Data for Figure 2, threshold trend plot for each stimulus condition
Thresholds were caluclted using Hamiltonian Monte Carlo estimation of three parametets, $\alpha, \beta, \lambda$, of this function:

<br>
$\psi(x; \alpha, \beta, \lambda) = 0.5 + (0.5 - \lambda)\Phi(x; \alpha, \beta)$, 
<br>

where $\psi$ is the proportion of correct answers, $x$ is the lag-lead ratio, LLR, in dB, $\Phi$ is the normal cumulative distribution function with location parameter $\alpha$ and scale parameter $\beta$, and $\lambda$ is the lapse-rate parameter (cf.  Kingdom & Prins, 2016, Ch. 4). A threshold was calculated by inverting the functiont to the find the LLR corresponding to 75 % correct responses.  


This parametrization of $\Phi$ was used: $\Phi(\frac{(x -\alpha)} \beta)$, so the 75 % threholds is given by  

<br>
$\psi^{-1}(0.75) = \alpha + \beta [\Phi^{-1}(\frac{0.25}{0.5 - \lambda})]$, 
<br>  

For this parameterization, $\alpha$ is the stimulus level [dB] of the function at its midpoint between the lower and upper asymptote, i.e., $\Phi^{-1}(0.75 - \lambda/2)$ (close to 75 % for small lapse rate $\lambda$), and  $\beta$ is the stimulus interval [dB] corresponding to a region around the midpoint (middle 38 % of the function between its lower and upper asymptote), sometimes called *width* or *spread* of the psychometric function.  

Weakly infomred priors were used:

+ $\alpha \sim Uniform(-60, 10)$, beacuse 10 dB was the theoretical maximum given the experimental design, and -60 would be inaudible based on prior testing.  
+ $\beta \sim Uniform(0, 40)$, beacuse the function should increase with stimulus levels (no negative values), and 40 dB represents a large part of the available stimulus range.
+ $\lambda \sim Beta(1, 29)$, because lapses are rare and this distribution has 95 % of its probability < 0.1, with increasingly higher probability as one approaches zero from the right. 

### STAN model
The STAN model was saved in a separate file called `psychomteric_function_revised.stan`. Here is the content of the file (m = $\alpha$, w = $\beta$ above, for no particular reason):

```{}
data {
int<lower=1> N;
vector[N] x;
int<lower=0,upper=1> y[N];
}
parameters {
real m;
real<lower=0> w;
real<lower=0> lambda;
}
model {
m ~ uniform(-60,10);
w ~ uniform(0,40);
lambda ~ beta(1,29);
for (i in 1:N)
{
y[i] ~ bernoulli(0.5 + (0.5-lambda)*Phi((x[i]-m)/w));
}
}
```


<br>
<br>

### Function moving() that calls Rstan and outputs threshold estimates. 

```{r}
moving <- function(tau = 1, startday = 1, iter = 1e4, warmup = 2e3) {
  g <- d[d$tau == tau & d$day >= startday &  d$day < (startday + 10), ]
  y <- as.integer(g$pc)
  x <- g$ratio
  N <- as.integer(length(y))

  model <- stan("psychometric_function_revised.stan", data = list(N = N, x = x, y = y),
              chains = 1, iter = iter, warmup = warmup)
  print(model)
  
  draws <- extract(model)
  m <- as.vector(draws$m)
  w <- as.vector(draws$w) 
  lapse <- as.vector(draws$lambda)
  
  # Threhold at 75 % correct responses (by inverting psychometric function)
  th <- m + w * qnorm(0.25/(0.5 - lapse))
  thhdi <- HPDI(th, prob = 0.95)
  names(thhdi) <- c('', '')
 
  out <- c(ici = tau, startday = startday, th = chainmode(th), 
           hdi95lo = thhdi[1], hdi95hi = thhdi[2], w = chainmode(w), lapse = chainmode(lapse))
  out
}
```
<br>
<br>

The function `moving()` was called separately for each of the 13 experimental conditions and each 10-day period (Day 1-10, 2-11, 51-60). In total $13 \times 51 = 663$ times. The script is below, BUT:
<br>
<br>
**WARNING! THIS WAS EXTREMELY SLOW** 
<br>
<br>
It took me about 36 hours. I had to do it two times, because my computer crashed half-way through. One reason it is slow is that the STAN-code is recompiled each time `stan()` is called (to avoid that R craches according to message) and that takes plenty of time. I am sure there is a way to specify a STAN-model that does everything in on step (the STAN manual give examples of moving-average models for time series), but I have not figured out how. 

```{r, eval = FALSE}

# Define conditions: startday x ici
cond <- expand.grid(startday = 1:51, ici = sort(unique(d$tau)))
all_data_revised <- data.frame(matrix(nrow = 0, ncol = 7))  # Empty data frame

# This has to be done twice, as R craches if you try all 600+ conditions at once
# Data stored in two files: all_data_stan_revised.RData, 
# which then are combined. 

# Get data, First from 1 to 320, ...
for (j in 1:320) {
  estimates <- moving(tau = cond[j, 2], startday = cond[j, 1])
  all_data_revised <- rbind(all_data_revised, estimates) # Combine with data from previous iteration
  
  # Save data, R may crach, so better save after each iteration
  save(all_data_revised, file = "all_data_stan_revised.RData")  
}

# and then fron 321 to nrow(cond)
for (j in 321:nrow(cond)){
  estimates <- moving(tau = cond[j, 2], startday = cond[j, 1])
  all_data_revised <- rbind(all_data_revised, estimates) # Combine with data from previous iteration
  
  # Save data, R may crach, so better save after each iteration
  save(all_data_revised, file = "all_data_stan_revised2.RData")  
}
```

<br>

### Store data for Figure 2
```{r, eval = FALSE}
# Data was stored in two files for reasons given above. This code loads the two files 
# and combines them into one datafram

load("all_data_stan_revised.RData")
d1 <- all_data_revised
colnames(d1) <- rep('', 7)  # Get rid of strange colnames

load("all_data_stan_revised2.RData")
d2 <- all_data_revised
colnames(d2) <- colnames(d2) <- rep('', 7)  # Get rid of strange colnames

standata_revised <- rbind(d1, d2)  # Combine data from the two runs

colnames(float_stan) <- c('ici', 'startday', 'thmode', 'thhdi95lo', 
                                'thhdi95hi', 'spreadmode', 'lapsemode')
# Save combined data to file
save(standata_revised, file = "standata.RData") 
```

The output was compiled in a file with 663 rows, one for each iteration of moving(). This data was stored in an R-object called `standata.RData`. It is available in the alexis108 package, or may be downloaded from
github.com/stamnosslin/alexis108/tree/master/data

<br>
<br>
<br>
<br>

## Data for Figure 3, thresholds first 10 versus last 10 days

### STAN model for comparing first and last 10 days
To compare thresholds from days 1-10 with days 51-60, a stan-model was called from a separate file called `prepost.stan`. Here is the content of the file:
```{}
data {
int<lower=1> N;
vector[N] x;
int<lower=0,upper=1> y[N];
int<lower=0,upper=1> z[N];
}
parameters {
real m1;
real<lower=0> w1;
real m2;
real<lower=0> w2;
real<lower=0, upper=1> lambda1;
real<lower=0, upper=1> lambda2;
}
model {
m1 ~ uniform(-60,10);
w1 ~ uniform(0,40);
m2 ~ uniform(-60,10);
w2 ~ uniform(0,40);
lambda1 ~ beta(1,29);
lambda2 ~ beta(1,29);
for (i in 1:N)
{
y[i] ~ bernoulli(0.5 + (0.5-((1-z[i])* lambda1 + z[i]*lambda2))*Phi((x[i]-((1-z[i])*m1 + z[i]*m2))/((1-z[i])*w1 + z[i]*w2)));
}
}
```

m1, w1, and lambda1 are location, scale and lapse rate for the first 10 days;  
m2, w2, and lambda2 are location, scale and lapse rate for the last 10 days, 

<br>

### Function that calls stan() and outputs estimates comparing data from Day 1-10 with data from Day 51-60

This fixes the data to dataframe g used as global variable by the functionfirstlast()
```{r}
d <- data108
d$pc <- 1 * (d$correct == d$answer)  # pc = 1 if answer was correct, 0 if not
g <- d[(d$day < 11 |  d$day > 50),]  # Remove everything except first and last 10 days
g$prepost <- (1 * g$day > 50)
```

<br>
Here is teh function:
```{r}
firstlast <- function(ici, iter = 1e4, warmup = 2e3) {
  y <- as.integer(g$pc[g$tau == ici])
  x <- g$ratio[g$tau == ici]
  z <- as.integer(g$prepost[g$tau == ici])
  N <- as.integer(length(y))
  model <- stan("prepost_revised.stan", data = list(N = N, x = x, y = y, z = z), 
              chains = 1, iter = iter, warmup = warmup)
  print(model)
  draws <- extract(model)
  m1 <- as.vector(draws$m1)
  w1 <- as.vector(draws$w1)
  lapse1 <- as.vector(draws$lambda1)
  m2 <- as.vector(draws$m2)
  w2 <- as.vector(draws$w2)
  lapse2 <- as.vector(draws$lambda2) 
  
  # Thresholds and 95 % HDI
  thpre <- m1 + w1 * qnorm(0.25/(0.5 - lapse1))
  hdipre <- HPDI(thpre, prob = 0.95)
  thpost <- m2 + w2 * qnorm(0.25/(0.5 - lapse2))
  hdipost <- HPDI(thpost, prob = 0.95)
  
  # Threshold difference
  thdiff <- thpost - thpre
  hdidiff <- HPDI(thdiff, prob = 0.95)
  
  # Diagnostics
  mineff <- min(summary(model)$summary[1:6, 9])  #Minimum n_eff 
  maxR <- max(summary(model)$summary[1:6, 10])  #Minimum n_eff 
  
  out <- c(ici = ici, 
           chainmode(thpre), hdipre, 
           chainmode(thpost), hdipost,
           chainmode(thdiff), hdidiff,
           chainmode(w1), chainmode(w2),
           chainmode(lapse1), chainmode(lapse2),
           mineff, maxR)
  out
}

```


<br>

### Store data for Figure 3
The output was compiled in a file with 26 rows, one for each iteration of `firstlast()`. This data was stored in an R-object called `prepost_stan.RData`. It is available in the alexis108 package, or it may be downloaded from github.com/stamnosslin/alexis108/tree/master/data
```{r, eval = FALSE}
conditions <- sort(unique(g$tau))
prepost_stan <- sapply(conditions, firstlast)  # Analysis conducted here
prepost_stan <- data.frame(t(prepost_stan)) # Reformat matrix
colnames(prepost_stan) <- c('ici', 'thmode_pre', 'hdi95pre_lo', 'hdi95pre_hi',
                                   'thmode_post', 'hdi95post_lo', 'hdi95post_hi', 
                                   'thdiff', 'hdi95diff_lo', 'hdi95diff_hi',
                                    'wmode_pre', 'wmode_post', 
                                    'lapsemdn_pre', 'lapsemdn_post',
                                    'min_neff', 'maxR')

save(prepost_stan, file = "prepost_stan.RData")
```


<br>
<br>
<br>
<br>
